---
title: "Seminar 6: Dimensionality reduction and Cluster Analysis-Hierarchical Clustering"
output: github_document
---

```{r setup, include=FALSE}
library(RColorBrewer)
library(cluster)
library(pvclust)
library(xtable)
library(limma)
library(plyr)
library(lattice)
library(RCurl)
options(download.file.method = "curl")
library(GEOquery)
library(knitr)
library(pheatmap)
library(ggplot2)
```


Read in data and do some cleaning:

```{r}
if (file.exists("GSE70213.Rdata")) {
    # if previously downloaded
    load("GSE70213.Rdata")
} else {
    # Get geo object that contains our data and phenotype information
    geo_obj <- getGEO("GSE70213", GSEMatrix = TRUE)
    geo_obj <- geo_obj[[1]]
    save(geo_obj, file = "GSE70213.Rdata")
}
# Get expression data
data <- exprs(geo_obj)

# Get covariate data
prDes <- pData(geo_obj)[, c("organism_ch1", "title", colnames(pData(geo_obj))[grep("characteristics", 
    colnames(pData(geo_obj)))])]

## Clean up covariate data
colnames(prDes) = c("organism", "sample_name", "tissue", "genotype", "sex", "age")
prDes$tissue = as.factor(gsub("tissue: ", "", prDes$tissue))
prDes$genotype = as.factor(gsub("genotype: ", "", prDes$genotype))
prDes$sex = as.factor(gsub("Sex: ", "", prDes$sex))
prDes$age = gsub("age: ", "", prDes$age)
```

Standardizing:

```{r}
sprDat <- t(scale(t(data)))
str(sprDat, max.level = 0, give.attr = FALSE)
round(data.frame(avgBefore = rowMeans(head(data)), avgAfter = rowMeans(head(sprDat)), 
    varBefore = apply(head(data), 1, var), varAfter = apply(head(sprDat), 1, var)), 
    2)
```

# Deliverable 1
## Regenerate the pheatmap clustering plot for the top genes, selected from limma, using clustering distance: correlation, and clustering method: mcquitty.

```{r}
# compute pairwise distances
data_to_plot = sprDat

pr.dis <- dist(t(data_to_plot), method = "euclidean")

# create a new factor representing the interaction of tissue type and genotype
prDes$grp <- with(prDes, interaction(tissue, genotype))

cutoff <- 1e-05
DesMat <- model.matrix(~grp, prDes)
dsFit <- lmFit(sprDat, DesMat)
dsEbFit <- eBayes(dsFit)
dsHits <- topTable(dsEbFit, coef = grep("grp", colnames(coef(dsEbFit))), 
                   p.value = cutoff, 
    n = Inf)
numBHhits <- nrow(dsHits)

topGenes <- rownames(dsHits)

# Scaled data of topGenes
topDat <- sprDat[topGenes, ]

clust_dist_col = "correlation"   

clust_method = "mcquitty"  

clust_scale = "none" 

pheatmap(topDat, cluster_rows = TRUE, scale = clust_scale,
         clustering_method = clust_method,
         clustering_distance_cols = clust_dist_col, 
         show_colnames = T, show_rownames = FALSE, 
         main = "Clustering heatmap for Top Genes selected from limma", 
         annotation = prDes[, c("tissue", "genotype", "grp")])

```

#Deliverable 2
## Regenerate the dendrogram on the SAMPLES of this heatmap using the 'hclust' and 'dist' functions.

K-means clustering:

```{r}
set.seed(31)
k <- 5
pr.km <- kmeans(topDat, centers = k, nstart = 50)
```

Generating the dendrogram:

```{r}
geneC.dis <- dist(topDat, method = "euclidean")

geneC.hc.a <- hclust(geneC.dis, method = "mcquitty")

plot(geneC.hc.a, labels = F, main = "Hierarchical with Average Linkage", xlab = "")
```

#Deliverable 3
## Plot the data for this analysis along PCs 1 and 2 using `ggplot` instead base plotting. Color the points by tissue.

```{r}
pcs <- prcomp(sprDat, center = FALSE, scale = FALSE)
prinComp <- cbind(prDes, pcs$rotation[rownames(prDes), 1:10])

ggplot(prinComp,aes(x=PC1,y=PC2,color=tissue)) + geom_point()
```


